--- 
title: "Survey Analysis"
author: "Patience Heath"
date: "2024-10-22"
output:
  html_document: default
  pdf_document: default
---

## Market Research Survey Analysis 
-- Group 6

We ended up conducting this survey using **Google Forms**

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(stringr)
survey_data = read_csv("AppForm.csv")
```


## Renaming columns to ease analysis 

A lot of the column names (survey questions) were long due to us wanting the user to understand each question. Although it was a little bit to decide what shorten names to use the most important thing is making sure that we get answers that are indicative of what the user feels. In order for us to do this we have to make sure the questions are clear and leave no room for confusion. 



```{r renaming columns}
new_survey_data = survey_data %>% rename(date= "Timestamp",
                                     age_range = `Select your age range:`,
                                     gender = `Select your gender:`,
                                     race = `Select your race:`,
                                     commute_or_residence = `Do you commute to campus, live on campus, or live off campus? (For college students only)`,
                                     living_area_description = `How would you describe the area you currently live?`,
                                     compare_before_purchase = `How often do you compare prices before making a purchase?`,
                                     most_interested_items = `Which types of items are you most interested in finding cheaper prices for? (Select all that apply)`,
                                     items_most_spent_monthly = `Which of the following items do you spend the most money on monthly? (select all that apply)`,
                                     price_comparison_methods = `Which of the following price comparison methods do you currently use? (Select all that apply)`,
                                     satisfaction_with_current_methods = `How satisfied are you with your current methods of finding the lowest prices?`,
                                     satisfaction_with_search_engines = `How satisfied are you with search engines’ results of locating the closest and cheapest options available to you?`,
                                     importance_in_saving = `How important is it for you to save money on everyday purchases?` ,
                                     shopping_methods = `How do you typically shop?` ,
                                     price_or_convenience = `Do you usually prioritize price or convenience when shopping?`,
                                     apps = `If possible, please list apps you use to find affordable items within your area.`)

```

## New Column names

```{r new column names}
colnames(new_survey_data)
```
For reproducibility I saved a separate file called column names with both the new and old column names. This is to be used for reference if someone is accessing the raw survey data.

```{r old and new column names}
colnames_new = colnames(new_survey_data)
colnames_old = colnames(survey_data)

names1 = as.data.frame(colnames_new, row.names = seq_along(colnames_new))
names2 = as.data.frame(colnames_old, row.names = seq_along(colnames_old))

survey_names_reference = cbind(names2,names1)
write.csv(survey_names_reference, file = "surveynames.csv")
```

We are going to take a first peak into the survey data to see what changes need to be made.

```{r glimpse survey data}
glimpse(new_survey_data)
```
## Changing the factor variables from character to factor

There a several factor variables that are just listed as character values. Listing these objects as character values diminishes the real meaning behind them. Converting to factor also makes it easier for us to make graphs of the data when the times comes to do so. Relationships are KEY and doing this helps of discover the relationships in the data.

```{r character to factor changes}
# unique(new_survey_data$age_range)

new_survey_data = new_survey_data %>% mutate(age_range = factor(age_range,
                                                        levels = c("Under 18",
                                                                   "18-24",
                                                                   "25-34",
                                                                   "35-44")))

# unique(new_survey_data$gender)

new_survey_data = new_survey_data %>% mutate(gender = factor(gender, 
                              levels = c("Male", "Female")))

# unique(new_survey_data$race)

new_survey_data = new_survey_data %>% mutate(race = factor(race, levels = 
c("Asian or Pacific Islander", "Hispanic Origin", "White", 
  "Black or African American", "Prefer not to say")))

# unique(new_survey_data$commute_or_residence)

new_survey_data = new_survey_data %>% mutate(commute_or_residence = 
                                       factor(commute_or_residence, levels = c(
                                         "Commuter", "Not a college student", 
                                         "Off Campus", "On Campus"
                                       )) )
# unique(new_survey_data$living_area_description)

new_survey_data = new_survey_data %>% mutate(living_area_description = factor(
  living_area_description, levels = c("Suburb", "City", "Rural area" )
))

# unique(new_survey_data$compare_before_purchase)

new_survey_data = new_survey_data %>% mutate(compare_before_purchase = factor(
  compare_before_purchase, levels = c("Rarely", "Sometimes","Often", "Always" )
))

# unique(new_survey_data$satisfaction_with_current_methods)

new_survey_data = new_survey_data %>% mutate(satisfaction_with_current_methods = 
                                       factor(satisfaction_with_current_methods,
                                              levels = c("1", "2", "3", "4", "5")))
# unique(new_survey_data$satisfaction_with_search_engines)

new_survey_data = new_survey_data %>% mutate(satisfaction_with_search_engines =
                                       factor(satisfaction_with_search_engines,
                                              levels = c(
                                                "1",
                                                "2",
                                                "3", 
                                                "4",
                                                "5"
                                              )))
# unique(new_survey_data$importance_in_saving)
new_survey_data = new_survey_data %>% mutate(importance_in_saving =
                                       factor(importance_in_saving,
                                              levels = c(
                                                "2",
                                                "3", 
                                                "4",
                                                "5"
                                              )))
# unique(new_survey_data$shopping_methods) did this while I worked to see unique values

new_survey_data = new_survey_data %>% mutate(shopping_methods = 
                                       factor(shopping_methods, levels = c(
                                         "A mix of online and in-store", 
                                         "Mostly in-store",
                                         "Mostly online"
                                       )))
#unique(survey_data$price_or_convenience) #there is some type of repeat here so
#this is going to be fixed 

#unique(survey_data$apps) #replace none with NA and then organize the rest of 
#these so that its just one word or something we will figure it out.
                                    
```

```{r second head check of new suvey data}
head(new_survey_data, n = 5)
```



```{r interest column}
#Game plan when it comes to this column.
#make a separate column for each category then dummy variable it.
#Example: Groceries 1, Gasoline 1, Clothing 1, health & beauty 0...
new_survey_data = new_survey_data %>% mutate(groceries_interest = if_else(
  str_detect(most_interested_items, "Groceries"), 1,0),
  clothing_interest = if_else(str_detect(most_interested_items, "Clothing"),1,0),
  electronics_interest = if_else(str_detect(
    most_interested_items, "Electronics"),1,0),
  household_items_interest = if_else(str_detect(
    most_interested_items, "Household items"),1,0),
  health_beauty_products_interest = if_else(str_detect(most_interested_items,
                                   "Health & beauty products"),1,0),
  entertainment_interest = if_else(str_detect(most_interested_items,
                                              "\\s*(?i)Entertainment"),1,0),
  gasoline_interest = if_else(str_detect(most_interested_items, 
                                         "\\s*(?i)Gasoline"), 1,0),
  other_interest = case_when(
     str_detect(most_interested_items, "Groceries") == FALSE &
     str_detect(most_interested_items, "Clothing") == FALSE &
     str_detect(most_interested_items, "Electronics") == FALSE &
     str_detect(most_interested_items, "Household items") == FALSE &
     str_detect(most_interested_items,"Health & beauty products") == FALSE &
     str_detect(most_interested_items,"Entertainment (movies, 
                                              events, etc..)") == FALSE &
     str_detect(most_interested_items, "Gasoline") == FALSE ~ 1,
     TRUE ~ 0
     ))

```

Checking as we go to make sure the code is correct.
**I manually checked the first ten rows to make sure that everything matched up correctly and it did so yay!**

```{r 3rd head check for sanity purposes}
head(survey_data, n =10)
```

Handling this column like we handled the previous
categorizing food as groceries , coffee as groceries , subscription app as 
entertainment but string might be able to detect that.


```{r items most spent monthly }


# unique(new_survey_data$items_most_spent_monthly) 


#might have duplicates but it doesn't really matter for the analysis 
#that I am doing 

new_survey_data = new_survey_data %>% mutate(items_most_spent_monthly = 
                                       str_replace(items_most_spent_monthly,
                                                   "(?i)food", "Groceries"))

new_survey_data = new_survey_data %>% mutate(items_most_spent_monthly = 
          str_replace(items_most_spent_monthly, "(?i)Coffee", "Groceries"))

## unique(new_survey_data$items_most_spent_monthly) 

new_survey_data = new_survey_data %>% mutate(groceries_spent_most = if_else(
  str_detect(items_most_spent_monthly, "\\s*(?i)Groceries"), 1,0),
  clothing_spent_most = if_else(str_detect(items_most_spent_monthly, "\\s*Clothing"),1,0),
  electronics_spent_most = if_else(str_detect(
    items_most_spent_monthly, "\\s*(?i)Electronics"),1,0),
  household_spent_most = if_else(str_detect(
    items_most_spent_monthly, "\\s*Household items"),1,0),
  Health_beauty_spent_most = if_else(str_detect(items_most_spent_monthly,
                                   "\\s*Health & beauty products"),1,0),
  entertainment_spent_most = if_else(str_detect(items_most_spent_monthly,
                                              "\\s*(?i)Entertainment"),1,0),
  gasoline_spent_most = if_else(str_detect(items_most_spent_monthly, 
                                        "\\s*Gasoline"), 1,0),
  other_spent_most = case_when(
     str_detect(items_most_spent_monthly, "Groceries") == FALSE &
     str_detect(items_most_spent_monthly, "Clothing") == FALSE &
     str_detect(items_most_spent_monthly, "Electronics") == FALSE &
     str_detect(items_most_spent_monthly, "Household items") == FALSE &
     str_detect(items_most_spent_monthly,"Health & beauty products") == FALSE &
     str_detect(items_most_spent_monthly,"Entertainment (movies, 
                                              events, etc..)") == FALSE &
     str_detect(items_most_spent_monthly, "Gasoline") == FALSE ~ 1,
     TRUE ~ 0
     ))

```

Checking as we go to make sure the code is correct.
I am not going to print this chunk but just wanted to show where my mind was at while doing each evaluation.

```{r, eval=FALSE}
head(new_survey_data, n = 10)
```


```{r price comparison methods}
# unique(new_survey_data$price_comparison_methods) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)


new_survey_data = new_survey_data %>% mutate(in_store_check = if_else(
  str_detect(price_comparison_methods, "In-store"),1,0),
  online_search_check = case_when(
    str_detect(price_comparison_methods, "\\s*(?i)online") |
      str_detect(price_comparison_methods,"\\s*google") ~ 1,
    TRUE ~ 0
  ),
  retailer_check =  if_else(str_detect(price_comparison_methods, "website"),1,0),
  word_of_mouth = if_else(str_detect(price_comparison_methods, "\\s*mouth"),1,0),
  price_app_check = if_else(str_detect(price_comparison_methods,
                                       "\\s*comparison apps"),1,0))
  
  
```

Same concept as above 
```{r, eval = FALSE}
head(new_survey_data, n = 10)
```


```{r price or convenience fix}
#unique(new_survey_data$price_or_convenience) #there is some type of repeat here so this is going to be fixed
#there was a repeat on the same value but case was different just ended up converting to sentence case

new_survey_data = new_survey_data %>% mutate(price_or_convenience = str_to_sentence(price_or_convenience)) 

unique(new_survey_data$price_or_convenience) #error fixed

```


```{r apps used}
unique(new_survey_data$apps)
new_survey_data = new_survey_data %>% mutate(no_apps_used = 
                      case_when(str_detect(apps,"(?i)\\s*N/?A\\s*") == TRUE |
                                 str_detect(apps,"(?i)\bdon'?t\b") == TRUE |
                                str_detect(apps,"\\.") == TRUE |
                              str_detect(apps, "(?i)\\s*instore") == TRUE |
                              str_detect(apps,"(?i)\\s*x") == TRUE |
                               str_detect(apps,"(?i)without") == TRUE ~ 1,
                            TRUE ~ 0),
      online_search_engine_used = case_when(
        str_detect(apps,"(?i)\\s*google") == TRUE |
           str_detect(apps,"(?i)\\s*safari") == TRUE |
          str_detect(apps, "(?i)\\s*online search") == TRUE ~ 1,
        TRUE ~ 0
      ),
            brand_website = if_else(str_detect(apps,"(?i)\\s*website"),1,0),
                      brand_apps = case_when(
                         str_detect(apps,"(?i)\\s*amazon") == TRUE |
                           str_detect(apps,"(?i)\\s*walmart") == TRUE |
                           str_detect(apps,"(?i)\\s*target") == TRUE |
                           str_detect(apps,"(?i)\\s*costco") == TRUE |
                           str_detect(apps,"(?i)\\s*aldis") == TRUE |
                           str_detect(apps,"(?i)\\s*trader") == TRUE |
                           str_detect(apps,"(?i)\\s*ross") == TRUE  |
                           str_detect(apps,"(?i)\\s*depop") == TRUE |
                           str_detect(apps, "(?i)\\s*ebay") == TRUE ~ 1,
                         TRUE ~ 0
                       ),
                       social_media = case_when(
                         str_detect(apps,"(?i)\\s*insta") == TRUE |
                           str_detect(apps,"(?i)\\s*facebook") == TRUE |
                           str_detect(apps,"(?i)\\s*whatsapp") == TRUE ~ 1,
                         TRUE ~0
                       ),
            gasbuddy = if_else(str_detect(apps, "(?i)\\s*gas"),1,0),
            honey = if_else(str_detect(apps, "(?i)\\s*honey"),1,0),
            retailmenot = if_else(str_detect(apps,"(?i)\\s*retailmenot"),1,0))

```


Checking as I go everything seems correct !!

```{r, eval = FALSE}
head(survey_data, n =10)
```


```{r}
glimpse(new_survey_data)
```

# Saving as a new csv and .Rdata

```{r}
write.csv(new_survey_data,file = "Group6.csv",row.names = FALSE)
saveRDS(new_survey_data, file = "survey_data.RData")

```



