--- 
title: "Survey Analysis"
author: "Patience Heath"
date: "2024-10-22"
output:
  html_document: default
  pdf_document: default
---

## Market Research Survey Analysis 
-- Group 6


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(stringr)
survey_data = read_csv("AppForm.csv")
```


```{r}
#pulling the column names 
survey_names = colnames(survey_data)
#converting it into a dataframe with column_names as the column name
survey_names_df = data.frame(column_number = seq_along(survey_names),
                             column_name = survey_names)
#writing CSV just of column names since I am renaming every row for convenience 
write.csv(survey_names_df, file = "survey_names.csv", row.names = FALSE)

```


```{r}
head(survey_data, n = 5)
```

## Renaming columns to ease analysis 


```{r}
survey_data = survey_data %>% rename(date= "Timestamp",
                                     age_range = `Select your age range:`,
                                     gender = `Select your gender:`,
                                     race = `Select your race:`,
                                     commute_or_residence = `Do you commute to campus, live on campus, or live off campus? (For college students only)`,
                                     living_area_description = `How would you describe the area you currently live?`,
                                     compare_before_purchase = `How often do you compare prices before making a purchase?`,
                                     most_interested_items = `Which types of items are you most interested in finding cheaper prices for? (Select all that apply)`,
                                     items_most_spent_monthly = `Which of the following items do you spend the most money on monthly? (select all that apply)`,
                                     price_comparison_methods = `Which of the following price comparison methods do you currently use? (Select all that apply)`,
                                     satisfaction_with_current_methods = `How satisfied are you with your current methods of finding the lowest prices?`,
                                     satisfaction_with_search_engines = `How satisfied are you with search engines’ results of locating the closest and cheapest options available to you?`,
                                     importance_in_saving = `How important is it for you to save money on everyday purchases?` ,
                                     shopping_methods = `How do you typically shop?` ,
                                     price_or_convenience = `Do you usually prioritize price or convenience when shopping?`,
                                     apps = `If possible, please list apps you use to find affordable items within your area.`)

```

New Column Names 

```{r}
colnames(survey_data)
```

```{r}
glimpse(survey_data)
```
## Changing the factor variables from character to factor

```{r}
unique(survey_data$age_range)

survey_data = survey_data %>% mutate(age_range = factor(age_range,
                                                        levels = c("Under 18",
                                                                   "18-24",
                                                                   "25-34",
                                                                   "35-44")))

unique(survey_data$gender)

survey_data = survey_data %>% mutate(gender = factor(gender, 
                              levels = c("Male", "Female")))

unique(survey_data$race)

survey_data = survey_data %>% mutate(race = factor(race, levels = 
c("Asian or Pacific Islander", "Hispanic Origin", "White", 
  "Black or African American", "Prefer not to say")))

unique(survey_data$commute_or_residence)

survey_data = survey_data %>% mutate(commute_or_residence = 
                                       factor(commute_or_residence, levels = c(
                                         "Commuter", "Not a college student", 
                                         "Off Campus", "On Campus"
                                       )) )
unique(survey_data$living_area_description)

survey_data = survey_data %>% mutate(living_area_description = factor(
  living_area_description, levels = c("Suburb", "City", "Rural area" )
))

unique(survey_data$compare_before_purchase)

survey_data = survey_data %>% mutate(compare_before_purchase = factor(
  compare_before_purchase, levels = c("Rarely", "Sometimes","Often", "Always" )
))

#unique(survey_data$most_interested_items) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)

#unique(survey_data$items_most_spent_monthly) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)

#unique(survey_data$price_comparison_methods) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)

unique(survey_data$satisfaction_with_current_methods)

survey_data = survey_data %>% mutate(satisfaction_with_current_methods = 
                                       factor(satisfaction_with_current_methods,
                                              levels = c("1", "2", "3", "4", "5")))
unique(survey_data$satisfaction_with_search_engines)

survey_data = survey_data %>% mutate(satisfaction_with_search_engines =
                                       factor(satisfaction_with_search_engines,
                                              levels = c(
                                                "1",
                                                "2",
                                                "3", 
                                                "4",
                                                "5"
                                              )))
unique(survey_data$importance_in_saving)
survey_data = survey_data %>% mutate(importance_in_saving =
                                       factor(importance_in_saving,
                                              levels = c(
                                                "2",
                                                "3", 
                                                "4",
                                                "5"
                                              )))
unique(survey_data$shopping_methods)

survey_data = survey_data %>% mutate(shopping_methods = 
                                       factor(shopping_methods, levels = c(
                                         "A mix of online and in-store", 
                                         "Mostly in-store",
                                         "Mostly online"
                                       )))
#unique(survey_data$price_or_convenience) #there is some type of repeat here so
#this is going to be fixed 
#unique(survey_data$apps) #replace none with NA and then organize the rest of 
#these so that its just one word or something we will figure it out.
                                    
```

```{r}
head(survey_data, n = 5)
```

Game plan when it comes to this column.

make a separate column for each category then dummy variable it.
Example: Groceries 1, Gasoline 1, Clothing 1, health & beauty 0...


```{r}
unique(survey_data$most_interested_items) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)

survey_data = survey_data %>% mutate(groceries_interest = if_else(
  str_detect(most_interested_items, "Groceries"), 1,0),
  clothing_interest = if_else(str_detect(most_interested_items, "Clothing"),1,0),
  electronics_interest = if_else(str_detect(
    most_interested_items, "Electronics"),1,0),
  household_items_interest = if_else(str_detect(
    most_interested_items, "Household items"),1,0),
  health_beauty_products_interest = if_else(str_detect(most_interested_items,
                                   "Health & beauty products"),1,0),
  entertainment_interest = if_else(str_detect(most_interested_items,
                                              "\\s*(?i)Entertainment"),1,0),
  gasoline_interest = if_else(str_detect(most_interested_items, 
                                         "\\s*(?i)Gasoline"), 1,0),
  other_interest = case_when(
     str_detect(most_interested_items, "Groceries") == FALSE &
     str_detect(most_interested_items, "Clothing") == FALSE &
     str_detect(most_interested_items, "Electronics") == FALSE &
     str_detect(most_interested_items, "Household items") == FALSE &
     str_detect(most_interested_items,"Health & beauty products") == FALSE &
     str_detect(most_interested_items,"Entertainment (movies, 
                                              events, etc..)") == FALSE &
     str_detect(most_interested_items, "Gasoline") == FALSE ~ 1,
     TRUE ~ 0
     ))
```
Checking as we go to make sure the code is correct.

```{r}
head(survey_data, n =10)
```

Handling this column like we handled the previous
categorizing food as groceries , coffee as groceries , subscription app as 
entertainment but string might be able to detect that.


```{r}


unique(survey_data$items_most_spent_monthly) 


#might have duplicates but it doesn't really matter for the analysis 
#that I am doing 

survey_data = survey_data %>% mutate(items_most_spent_monthly = 
                                       str_replace(items_most_spent_monthly,
                                                   "(?i)food", "Groceries"))

survey_data = survey_data %>% mutate(items_most_spent_monthly = 
          str_replace(items_most_spent_monthly, "(?i)Coffee", "Groceries"))

unique(survey_data$items_most_spent_monthly) 

survey_data = survey_data %>% mutate(groceries_spent_most = if_else(
  str_detect(items_most_spent_monthly, "\\s*(?i)Groceries"), 1,0),
  clothing_spent_most = if_else(str_detect(items_most_spent_monthly, "\\s*Clothing"),1,0),
  electronics_spent_most = if_else(str_detect(
    items_most_spent_monthly, "\\s*(?i)Electronics"),1,0),
  household_spent_most = if_else(str_detect(
    items_most_spent_monthly, "\\s*Household items"),1,0),
  Health_beauty_spent_most = if_else(str_detect(items_most_spent_monthly,
                                   "\\s*Health & beauty products"),1,0),
  entertainment_spent_most = if_else(str_detect(items_most_spent_monthly,
                                              "\\s*(?i)Entertainment"),1,0),
  gasoline_spent_most = if_else(str_detect(items_most_spent_monthly, 
                                        "\\s*Gasoline"), 1,0),
  other_spent_most = case_when(
     str_detect(items_most_spent_monthly, "Groceries") == FALSE &
     str_detect(items_most_spent_monthly, "Clothing") == FALSE &
     str_detect(items_most_spent_monthly, "Electronics") == FALSE &
     str_detect(items_most_spent_monthly, "Household items") == FALSE &
     str_detect(items_most_spent_monthly,"Health & beauty products") == FALSE &
     str_detect(items_most_spent_monthly,"Entertainment (movies, 
                                              events, etc..)") == FALSE &
     str_detect(items_most_spent_monthly, "Gasoline") == FALSE ~ 1,
     TRUE ~ 0
     ))

```

Checking as we go to make sure the code is correct.

```{r}
head(survey_data, n = 10)
```


```{r}
unique(survey_data$price_comparison_methods) # this will have to be handled 
#separately (probably will do a string operation then mutate another column on)


survey_data = survey_data %>% mutate(in_store_check = if_else(
  str_detect(price_comparison_methods, "In-store"),1,0),
  online_search_check = case_when(
    str_detect(price_comparison_methods, "\\s*(?i)online") |
      str_detect(price_comparison_methods,"\\s*google") ~ 1,
    TRUE ~ 0
  ),
  retailer_check =  if_else(str_detect(price_comparison_methods, "website"),1,0),
  word_of_mouth = if_else(str_detect(price_comparison_methods, "\\s*mouth"),1,0),
  price_app_check = if_else(str_detect(price_comparison_methods,
                                       "\\s*comparison apps"),1,0))
  
  
```

```{r}
head(survey_data, n = 10)
```


```{r}
unique(survey_data$price_or_convenience) #there is some type of repeat here so
#this is going to be fixed
#there was a repeat on the same value but case was different just ended up converting to sentence case

survey_data = survey_data %>% mutate(price_or_convenience = str_to_sentence(price_or_convenience)) 

unique(survey_data$price_or_convenience) #error fixed

```


```{r}
unique(survey_data$apps)

survey_data = survey_data %>% mutate(no_apps_used = 
                      case_when(str_detect(apps,"(?i)\\s*N/?A\\s*") == TRUE |
                                 str_detect(apps,"(?i)\bdon'?t\b") == TRUE |
                                str_detect(apps,"\\.") == TRUE |
                              str_detect(apps, "(?i)\\s*instore") == TRUE |
                              str_detect(apps,"(?i)\\s*x") == TRUE |
                               str_detect(apps,"(?i)without") == TRUE ~ 1,
                            TRUE ~ 0),
      online_search_engine_used = case_when(
        str_detect(apps,"(?i)\\s*google") == TRUE |
           str_detect(apps,"(?i)\\s*safari") == TRUE |
          str_detect(apps, "(?i)\\s*online search") == TRUE ~ 1,
        TRUE ~ 0
      ),
            brand_website = if_else(str_detect(apps,"(?i)\\s*website"),1,0),
                      brand_apps = case_when(
                         str_detect(apps,"(?i)\\s*amazon") == TRUE |
                           str_detect(apps,"(?i)\\s*walmart") == TRUE |
                           str_detect(apps,"(?i)\\s*target") == TRUE |
                           str_detect(apps,"(?i)\\s*costco") == TRUE |
                           str_detect(apps,"(?i)\\s*aldis") == TRUE |
                           str_detect(apps,"(?i)\\s*trader") == TRUE |
                           str_detect(apps,"(?i)\\s*ross") == TRUE  |
                           str_detect(apps,"(?i)\\s*depop") == TRUE |
                           str_detect(apps, "(?i)\\s*ebay") == TRUE ~ 1,
                         TRUE ~ 0
                       ),
                       social_media = case_when(
                         str_detect(apps,"(?i)\\s*insta") == TRUE |
                           str_detect(apps,"(?i)\\s*facebook") == TRUE |
                           str_detect(apps,"(?i)\\s*whatsapp") == TRUE ~ 1,
                         TRUE ~0
                       ),
            gasbuddy = if_else(str_detect(apps, "(?i)\\s*gas"),1,0),
            honey = if_else(str_detect(apps, "(?i)\\s*honey"),1,0),
            retailmenot = if_else(str_detect(apps,"(?i)\\s*retailmenot"),1,0))

```


Checking as I go everything seems correct !!
```{r}
head(survey_data, n =10)
```

## New column names 

```{r}
glimpse(survey_data)
```
```{r}
for (i in 17:ncol(survey_data)) {
  survey_data[[i]] = as.factor(survey_data[[i]])
}
```

```{r}
#glimpse(survey_data)
survey_data = survey_data %>% filter(!is.na(gasbuddy))
```

#Saving as a new csv and .Rdata

```{r}
write.csv(survey_data,file = "Group6.csv",row.names = FALSE)
saveRDS(survey_data, file = "survey_data.RData")

```



